A Systematic Literature Review on Role-Based Access Control

### 2013-02-20 paper/sections/analysis.tex
Cleans up outlines, adds numerous updates to various sections.
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

@@

When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, bringing the model to life is strong evidence that the
proposed model can work in practice.  The concept of authorization, and access control
is rooted in a business need.  Thus, any access control model needs to be feasible
in the real world not just on paper.  We analyzed the primary sources to see how many
proposed models actually had implmenetations associated with them.  And quantified the
type of implementation.  Whether the implementation was for a real system, for a prototype
and/or used in a production environment.

Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
mention of a implementation or prototype.  Of the remianing 8 papers that did mention an implementation, half 
were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
system.

@@

Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
and analsysis as a way to provide evaluation in the form of completeness.

@@

\subsection{Research Question X}

@@

>>>>

\subsection{Research Question 1}

What are the deficiencies in the current RBAC model? \\

Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
security mechanism, a variety of extended RBAC models have been proposed over the years to meet
security requirements. Our surveyed 29 papers address more than one deficiencies described as follows.

\begin{itemize}
%	\item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
	\item Core RBAC does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
	\item Core RBAC does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
	\item Core RBAC does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
	\item Core RBAC has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
\end{itemize}

\subsection{Research Question 2}

What are the motivations behind RBAC extensions? \\

When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
Administrators improve existing core RBAC model due to following reasons:

\begin{itemize}
	\item Model needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
	\item Model needs to incorporate additional attributes such as
	task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new association between attributes
	are introduced. 
		\item Model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
\end{itemize}

\subsection{Research Question 3}

@@

When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, bringing the model to life is strong evidence that the
proposed model can work in practice.  The concept of authorization, and access control
is rooted in a business need.  Thus, any access control model needs to be feasible
in the real world not just on paper.  We analyzed the primary sources to see how many
proposed models actually had implementations associated with them.  And quantified the
type of implementation.  Whether the implementation was for a real system, for a prototype
and/or used in a production environment.

Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
mention of a implementation or prototype.  Of the remaining 8 papers that did mention an implementation, half 
were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
system.

\subsection{Research Question 4}

@@

Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
and analysis as a way to provide evaluation in the form of completeness.

@@

@@

What commonalities or generalizations exist across all categorizations? \\

### 2013-02-20 paper/sections/intro.tex
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

Role based access control (RBAC) was first introduced in the 1990s when the National Instittute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.  In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions of specific actions on resources to roles instead of individual users.  Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.  RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.

@@

\textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}

A systematic literature review is designed to yield quanitfiable results across a body of research in order to draw out similarities, trends and defiencies in an area as a whole.  The review we performed yielded 1716, of which XX were deemed primary sources for inclusion as model extensions to the RBAC standard model.  This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel. For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.  Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.

>>>>

Role based access control (RBAC) was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.  In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions of specific actions on resources to roles instead of individual users.  Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.  RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.

@@

\textit{The goal of this work is to provide practitioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}

A systematic literature review is designed to yield quantifiable results across a body of research in order to draw out similarities, trends and deficiencies in an area as a whole.  The review we performed yielded 1716, of which XX were deemed primary sources for inclusion as model extensions to the RBAC standard model.  This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel. For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.  Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.

### 2013-02-20 paper/sections/results.tex
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

\begin{table}
\centering
\begin{tabular}{|p{12.5cm}|p{3cm}|}
\hline
\textbf{Paper} & \textbf{Category}
\\\hline
Alam, M. and Hafner, M. and Breu, R. 2006 \cite{alam06:constraint} & Constraint \\\hline
Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 \cite{tzelepi01:flexible} & Context \\\hline
Haibo, SHEN and Fan, HONG 2005 \cite{haibo05:context} & Context \\\hline
Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 \cite{huang06:pervasive} & Context \\\hline
Motta, G.H.M.B. and Furuie, S.S. 2003 \cite{motta03:contextual} & Context \\\hline
Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 \cite{bao08:role} & Context \\\hline
Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 \cite{yamazaki104:designing} & Context \\\hline
Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 \cite{han08:extended} & Context \\\hline
Thein, N. and others 2011 \cite{thein2011leveraging} & Context \\\hline
Zou, D. and He, L. and Jin, H. and Chen, X. 2009 \cite{zou2009crbac} & Context \\\hline
Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 \cite{hasebe10:capability} & Delegation \\\hline
Zhang, Z. and Zhang, X. and Sandhu, R. 2006 \cite{zhang06:collaborative} & Organizational \\\hline
Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 \cite{ni2010privacy} & Privacy \\\hline
Masoumzadeh, A. and Joshi, J. 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
Zhao, Y. and Zhao, Y. and Lu, H. 2008 \cite{zhao2008flexible} & Resource \\\hline
Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 \cite{damian2007geo} & Spatial \\\hline
Ray, I. and Kumar, M. and Yu, L. 2006 \cite{ray07:spatio} & Spatial \\\hline
Hansen, F. and Oleshchuk, V. 2003 \cite{hansen2003spatial} & Spatial \\\hline
Aich, S. and Sural, S. and Majumdar, A. 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
Chen, L. and Crampton, J. 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
Samuel, A. and Ghafoor, A. and Bertino, E. 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
Chandran, S. and Joshi, J. 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
Ray, I. and Toahchoodee, M. 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
ZHANG, S. and CHEN, X. and HOU, G. 2009 \cite{zhou2007team} & Task \\\hline
Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline

@@

\begin{itemize}
%	\item Access Control Model:
%	\item Access Control Policy:
%	\item Access Control Requirements:
%	\item Rules:
%	\item Attributes-Based Access Control:
	\item Task-Based Access Control:
%	\item Purpose-Based Access Control:
	\item Agent-Based Access Control:						
\end{itemize}

@@

@@

	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$. SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$. (Role-Based Access Control by F. Ferrailolo et al.)

	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user. (Role-Based Access Control by F. Ferrailolo et al.)
	\end{itemize}

@@

\begin{itemize}
	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.   
	(Role-Based Access Control by F. Ferrailolo et al.)

>>>>

\begin{table}
\centering
\begin{tabular}{|p{12.5cm}|p{3cm}|}
\hline
\textbf{Paper} & \textbf{Category}
\\\hline
Alam, M. and Hafner, M. and Breu, R. 2006 \cite{alam06:constraint} & Constraint \\\hline
Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 \cite{tzelepi01:flexible} & Context \\\hline
Haibo, SHEN and Fan, HONG 2005 \cite{haibo05:context} & Context \\\hline
Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 \cite{huang06:pervasive} & Context \\\hline
Motta, G.H.M.B. and Furuie, S.S. 2003 \cite{motta03:contextual} & Context \\\hline
Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 \cite{bao08:role} & Context \\\hline
Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 \cite{yamazaki104:designing} & Context \\\hline
Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 \cite{han08:extended} & Context \\\hline
Thein, N. and others 2011 \cite{thein2011leveraging} & Context \\\hline
Zou, D. and He, L. and Jin, H. and Chen, X. 2009 \cite{zou2009crbac} & Context \\\hline
Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 \cite{hasebe10:capability} & Delegation \\\hline
Zhang, Z. and Zhang, X. and Sandhu, R. 2006 \cite{zhang06:collaborative} & Organizational \\\hline
Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 \cite{ni2010privacy} & Privacy \\\hline
Masoumzadeh, A. and Joshi, J. 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
Zhao, Y. and Zhao, Y. and Lu, H. 2008 \cite{zhao2008flexible} & Resource \\\hline
Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 \cite{damian2007geo} & Spatial \\\hline
Ray, I. and Kumar, M. and Yu, L. 2006 \cite{ray07:spatio} & Spatial \\\hline
Hansen, F. and Oleshchuk, V. 2003 \cite{hansen2003spatial} & Spatial \\\hline
Aich, S. and Sural, S. and Majumdar, A. 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
Chen, L. and Crampton, J. 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
Samuel, A. and Ghafoor, A. and Bertino, E. 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
Chandran, S. and Joshi, J. 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
Ray, I. and Toahchoodee, M. 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
Zhang, S. and Chen, X. and Hou, G. 2009 \cite{zhou2007team} & Task \\\hline
Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline

@@

\begin{itemize}
%	\item Access Control Model:
%	\item Access Control Policy:
%	\item Access Control Requirements:
%	\item Rules:
%	\item Attributes-Based Access Control:

%http://www.hpi.uni-potsdam.de/fileadmin/hpi/Forschung/Publikationen/Dissertationen/Diss_Zhou.pdf
% A Task-Role Based Access Control Model with Multi-Constraints
	\item Task-role-based access control: a task is a fundamental unit of a business activity. Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions. Roles are first associated with tasks, which are associated permissions. For example, the employee role is associated with a task, which is to write a report. Then, this task is associated with a permission.

% Team and Task Based RBAC Access Control Model	
	\item Team-based access control: a team encapsulates a group of users, who have various roles. A team member can activate only roles that can be assigned to a team. Moreover, team-based access control may decide maximum number of permissions such as roles (that are assigned to a team) having a maximum number of permissions to the team.			
%	\item Purpose-Based Access Control:
%	\item Agent-based access control:						
\end{itemize}

@@

	\item Delegation: delegation is to assign permissions (to access specific resources) available to a user to another user.
	A employee $E_1$ may delegate her/his permissions to another employee $E_2$ while $E_1$ is on leave.

%	 obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
%	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
%	her/his office duty before accessing the resources.

@@

	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$. SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.

	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.
	\end{itemize}

@@

\begin{itemize}
	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
